---
title: "NEWS"
author: Erik.Leppo@tetratech.com
output: github_document
  #md_document:
   # variant: markdown_github
---

<!-- NEWS.md is generated from NEWS.Rmd. Please edit that file -->

```{r Options, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.path = "NEWS-")
```

```{r LastUpdate, echo = FALSE}
cat(paste0("Last Update: ",Sys.time()))
```

# Planned Updates

* Metric Calculation
* Subsample (rarify)
* Excluded Taxa
* TaxaList Check

# Future Updates

* Fish and Periphyton metrics.

# Version History

## Changes in Version 0.1.0.9002 (2018-11-16)

* extdata\Data_Benthos.xlsx
    + OneDrive messed up version included in package.

## Changes in Version 0.1.0.9001 (2018-11-16)

* Added basic package structure.
    + Folders (data, data-raw, inst/doc, inst/extdata, man, R, vignettes)
    + NEWS, README, LICENSE
* Update DESCRIPTION, NEWS, and README
* Add functions and data
    + rarify
    + metric.values

## Changes in Version 0.1.0 (2018-11-16)

* Initial commit on GitHub.

