---
title: "NEWS"
author: Erik.Leppo@tetratech.com
output: github_document
  #md_document:
   # variant: markdown_github
---

<!-- NEWS.md is generated from NEWS.Rmd. Please edit that file -->

```{r Options, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.path = "NEWS-")
```

```{r LastUpdate, echo = FALSE}
cat(paste0("Last Update: ",Sys.time()))
```

# Planned Updates

* ~~Metric Calculation~~
* ~~Subsample (rarify)~~
* ~~Excluded Taxa~~

# Future Updates

* Fish and Periphyton metrics.
* Metric scoring
* Multimetric Index scoring
* TaxaList Check
* Map taxon observations

# Version History

## Changes in Version 0.2.0.9005 (2018-11-26)

* Added MapTaxaObs.

## Changes in Version 0.2.0.9004 (2018-11-26)

* Percent metrics as 0-1.  Issue #9.

## Changes in Version 0.2.0.9003 (2018-11-26)

* Add qc.checks function from BCGcalc package.  Issue #9
    + qc.checks.R
    + Vignette
    + extdata\MetricFlags.xlsx
* Update read_excel with guess_max=10^6 for bio data file.

## Changes in Version 0.2.0.9002 (2018-11-20)

* Update SuperFamily Tipuloidea.
    + Update data and examples in function and vignette

## Changes in Version 0.2.0.9001 (2018-11-20)

* Add Travis CI.

## Changes in Version 0.2.0 (2018-11-20)

* Initial release version.

## Changes in Version 0.1.0.9004 (2018-11-20)

* Add SuperFamily
    + function
    + vignette
    + example data file
* Add metric names files to extdata and doc

## Changes in Version 0.1.0.9003 (2018-11-20)

* markExcluded
    + Add function with examples
* extdata\Data_Benthos.xlsx
    + Updated bad Excluded entries.
        - FALSE to TRUE, n=6
        - TRUE to FALSE, n=1
* DESCRIPTION
    + Update Imports and Suggests
* NEWS
    + Update Planned and Future Updates
* README
    + Tweak language in Usage section.
    + Update icons.
    + Add Travis CI.
* Vignette

## Changes in Version 0.1.0.9002 (2018-11-16)

* extdata\Data_Benthos.xlsx
    + OneDrive messed up version included in package.

## Changes in Version 0.1.0.9001 (2018-11-16)

* Added basic package structure.
    + Folders (data, data-raw, inst/doc, inst/extdata, man, R, vignettes)
    + NEWS, README, LICENSE
* Update DESCRIPTION, NEWS, and README
* Add functions and data
    + rarify
    + metric.values

## Changes in Version 0.1.0 (2018-11-16)

* Initial commit on GitHub.

