---
title: "NEWS"
author: Erik.Leppo@tetratech.com
output: github_document
  #md_document:
   # variant: markdown_github
---

<!-- NEWS.md is generated from NEWS.Rmd. Please edit that file -->

```{r Options, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.path = "NEWS-")
```

```{r LastUpdate, echo = FALSE}
cat(paste0("Last Update: ",Sys.time()))
```

# Planned Updates

* None at this time.

# Future Possible Updates

* Periphyton metrics.
* Metric scoring
* TaxaList Check
* Map taxon observations

# Version History

## Changes in Version 0.4.0.9001 (2020-02-27)

* Tweak for GA Fish IBI, Issue #3
    + Modify ni_natnonhybrid and ni_natnonhybridnonlepomis to also exclude mosquitofish.
        - MetricScoring.xlsx
        - metric.score.R
    + Add nt_nativenonhybrid
        - MetricScoring.xlsx
        - metric.score.R
* metric.score.R
    + Update ni_natnonhybridnonlepomis.

## Changes in Version 0.4.0 (2020-02-24)

* Changes to fish metric significant enough for a minor version update.
    + Some previous calculations may not work.  Function made more generic.

## Changes in Version 0.3.3.9017 (2020-02-24)

* MetricScoring.xlsx, Issue #3
    + Tweak scoring for GA Fish IBI
    + Add metric value calculation for GA Fish IBI metrics.
    + New Scoring Regimes and associated columns.
* Remove MBSStools from Suggests in DESCRIPTION.
* metric.values
    + Modifications for GA Fish IBI.
        - All columns to upper case.
    + Default values to NA rather than zero.
    + Use toupper rater than tolower.
* metric.scores
    + Modifications for GA Fish IBI.
        - Additional Scoring Regimes.
        - Additional fish metrics.
    + All columns to upper case using toupper.
    + Default values to NA rather than zero.
* vignette_BioMonTools.Rmd
    + Added library(dplyr) to one of the examples.

## Changes in Version 0.3.3.9016 (2020-01-31)

* MetricScoring.xlsx, Issue #3
    + Tweak values for GA Fish IBI.
        - ACF, 6b, inflection point is 2 not 1.

## Changes in Version 0.3.3.9015 (2020-01-08)

* metric.values.R
    + Minor edits.
* MetricScoring.xlsx, Issue #3
    + Tweak values for GA Fish IBI.
        - Convert some values to text.

## Changes in Version 0.3.3.9014 (2020-01-07)

* metric.values.R
    + Tweak index NormDist_135.
* MetricScoring.xlsx, Issue #3
    + Tweak values for GA Fish IBI.
        - Convert some values to text.

## Changes in Version 0.3.3.9013 (2020-01-07)

* metric.values.R
    + Tweak index sum to account for NA.
* MetricScoring.xlsx, Issue #3
    + Tweak values for GA Fish IBI.

## Changes in Version 0.3.3.9012 (2020-01-02)

* metric.values.R
    + Error in excluding marine metrics when no metric list given (i.e., want all metrics).
        - Misplaced ")".  Issue # 31.

## Changes in Version 0.3.3.9011 (2019-12-18)

* MetricScoring.xlsx, Issue #3, Issue #30
    + Added more columns for new scoring regimes for GA Fish IBI.
* MetricNames.xlsx
    + Added new fish metric names.
* metric.scores.R
    + Added more scoring regimes (GA Fish IBI).

## Changes in Version 0.3.3.9010 (2019-10-17)

* MetricScoring.xlsx, Issue #3
    + Numeric to text.

## Changes in Version 0.3.3.9009 (2019-10-17)

* MetricScoring.xlsx, Issue #3
    + Updates for GBI_MS_2013

## Changes in Version 0.3.3.9008 (2019-10-16)

* MetricScoring.xlsx, Issue #3
    + Updates for GBI_MS_2013
    + Numeric to text. 

## Changes in Version 0.3.3.9007 (2019-10-16)

* MetricScoring.xlsx, Issue #3
    + Updates for GBI_MS_2013
    + Numeric to text. 

## Changes in Version 0.3.3.9006 (2019-10-15)

* MetricScoring.xlsx, Issue #3
    + Updates for GBI_MS_2013

## Changes in Version 0.3.3.9005 (2019-10-15)

* MetricScoring.xlsx, Issue #3
    + Updates for GBI_MS_2013

## Changes in Version 0.3.3.9004 (2019-10-15)

* Additional scoring and index.  Issue #3
    + Gulf Biotic Index, 2011
    + Mississippi Gulf Biotic Index, 2013

## Changes in Version 0.3.3.9003 (2019-09-15)

* metric.values, Issue #3
   + Additional estuary/marine metrics
       - Gulf Biotic Index, 2011
       - Mississippi Gulf Biotic Index, 2013
   + Additional input parameter, boo.Marine
   + Data_Benthos.xlsx
      - Add "InfraOrder" column.
      - Require in metric.values

## Changes in Version 0.3.3.9002 (2019-07-03)

* metric.values, Issue #3
    + Add placeholders for marine/estuarine metrics
        - Not yet in MetricScoring.xlsx or MetricNames.xlsx

## Changes in Version 0.3.3.9001 (2019-07-02)

* metric.values
    + Update Beck's Biotic Index cutoff signs.
        - < 1.5 and >= 1.5 to <= 1.5 and > 1.5.
        - Only affects tolerance values of exactly 1.5.

## Changes in Version 0.3.3 (2019-07-02)

* New release version.

## Changes in Version 0.3.2.9001 (2019-07-02)

* metric.scoring.xlsx
    + Update "scoring formula" column for 2 MA WestHighlands metrics.
    + Actual thresholds not affected.

## Changes in Version 0.3.2 (2019-06-28)

* New release version.

## Changes in Version 0.3.1.9001 (2019-06-27)

* MetricScoring.xlsx, issue #22
    + New metric misspelled.
* Add capability for 6 narrative categories.
    + MetricScoring.xlsx
    + metric.scores.R
       - Minor edit for number of columns.

## Changes in Version 0.3.1 (2019-06-27)

* New release version.

## Changes in Version 0.3.0.9024 (2019-06-27)

* MetricScoring.xlsx, issue #22
    + Some numbers to character with ' prefix.
    + These numbers not importing correctly from Excel.
        - Floating point errors in rounding.

## Changes in Version 0.3.0.9023 (2019-06-27)

* MetricNames.xlsx
    + Add fish metric names, Issue #21
    + Demonstration only.  Not final format in function.
* MetricScoring.xlsx
    + Update date on Notes worksheet.
* metric.values.R
    + Update demonstration fish metrics, Issue #21.

## Changes in Version 0.3.0.9022 (2019-06-27)

* Check
    + DESCRIPTION
        - rmarkdown listed under imports and suggests.  Remove from suggests.
        - Add to suggests; ggplot2 and reshape2.
        - Add MBSStools to suggests.  May import the data later and remove.
    + Example, move "View" to dontrun. 
        - metric.scores.R
        - metric.values.R
        - qc.checks.R
        - rarify.R
    + Fix line widths greater than 100 characters:
        - markExcluded.R
        - metric.scores.R
    + Properly declare functions from other packages:
        - "grDevices", "dev.off", "jpeg", "pdf"
        - "graphics", "mtext", "points"
        - "stats", "median", "runif"
        - "utils", "flush.console"
    + MapTaxaObs.R
        - Document the "..." argument.

## Changes in Version 0.3.0.9021 (2019-06-27)

* Updated metric names for MA index, Issue #3
    + MetricScoring.xlsx
* Added "to do" and "references" to metricscoring.xlsx, Issue #3

## Changes in Version 0.3.0.9020 (2019-05-24)

* Updated metric names for MBISQ index, Issue #3
    + MetricScoring.xlsx

## Changes in Version 0.3.0.9019 (2019-05-24)

* Added metrics, Issue #3
    + metric.values
        - x_NCBI, x_D_Mg, x_D_G
        - Additional QC for TolVal2 as numeric (same as TolVal).
    + MetricNames.xlsx

## Changes in Version 0.3.0.9018 (2019-05-17)

* Fixes for R v3.6.0, Issue #18
    + DESCRIPTION
        - Remove StagedInstall: no
    + README
        - Add directions for fix for devtools.
* README
    + Update badges.

## Changes in Version 0.3.0.9017 (2019-05-08)

* Metric names, fix those added in previous update, Issue #3
    + Files
        - metric.values.R
        - MetricNames.xlsx
        - MetricScoring.xlsx
* metric.scores, Issue #19
    + Continuous 0-100 scoring only reporting 1 value per column.
    + Fixed so reports all values.
        - Was taking median of column not per row.

## Changes in Version 0.3.0.9016 (2019-05-07)

* Temp fix for staged install with R v3.6.0, Issue #18
    + DESCRIPTION, StagedInstall: no
* Add additional metrics (GA DNR).  Issue #3.
    + Files
        - metric.values.R
        - MetricNames.xlsx
        - MetricScoring.xlsx
    + Metrics
        - pi_Hydro2EPT
        - pi_Hydro2Trich
        - pi_Ortho2Chi
        - pi_Tanyp2Chi
        - pi_ChCr2Chi

## Changes in Version 0.3.0.9015 (2019-04-11)

* inst/extdata/MetricFlags.xlsx
   + Update ni_total for Hi to match Lo (both 450), Issue #16

## Changes in Version 0.3.0.9014 (2019-04-02)

* metric.values, Issue #15
    + Fix metric pi_EphemNoCaeBae
        - Was calculating the same as pi_Ephem.
    + Checked all other "No" metrics.
        - All ok.

## Changes in Version 0.3.0.9013 (2019-01-25)

* metric.values
    + Fix Trombidiformes metric.
        - Mispelled search term so wasn't working.
    + Added Megaloptera metrics (nt, pi, and pt).

## Changes in Version 0.3.0.9012 (2019-01-17)

* metric.values
    + Remove metric sorting.  Rely on fun.MetricNames to request and sort metrics.
    + Modify TolVal character check for NA.
    + Add example keeping only some metrics.
* extdata/MetricScoring.xlsx
    + Modify percent metric thresholds (from 0-1 to 0-100).
* extdata/MetricFlags.xlsx
    + Modify percent metric thresholds (from 0-1 to 0-100).

## Changes in Version 0.3.0.9011 (2019-01-17)

* docs/ExcludedTaxaDecisionCrit.pdf
    + Recreate.  Wasn't opening.

## Changes in Version 0.3.0.9010 (2019-01-17)

* extdata/MetricNames.xlsx
    + move "0-100" from notes to description.

## Changes in Version 0.3.0.9009 (2019-01-17)

* metric.values
    + Mistyped "FAMILY" as "Family" when added pi_Baet.

## Changes in Version 0.3.0.9008 (2019-01-17)

* docs
    + Added ExcludedTaxaDecisionCrit.docx

## Changes in Version 0.3.0.9007 (2019-01-17)

* Remove metric names PDF
* metric.scores
    + Add pi_Baet
    + Percent metrics to 0-100 from 0-1.
    + Add pi_Baet to extdata/MetricNames.xlsx
    + input TolVal "NA" (character) to NA.

## Changes in Version 0.3.0.9006 (2019-01-17)

* metric.scores
    + Additional metrics.
    + Metric name sort variable.  Issue #13
    + Organize metrics ni, nt, pi, pt for each section.
    + Update extdata/MetricNames.xlsx
* MapTaxaObs
    + Set up for ability to sort taxa but not active.

## Changes in Version 0.3.0.9005 (2019-01-14)

* DESCRIPTION
    + maps package missing from Imports.
        - Used in MapTaxaObs function.

## Changes in Version 0.3.0.9004 (2019-01-14)

* README
    + Update install example to force vignettes.
* metric.scores
    + pt_habit metrics (n=5) used ni_total instead of nt_total in denominator.

## Changes in Version 0.3.0.9003 (2019-01-14)

* MapTaxaObs
    + Account for tibbles as input.
        - Convert to a data frame inside the function.  Otherwise fails.
    + Subset function not working in every case.  Update line 128.
    + Add map inputs database and regions and funcion inputs.
* extdata/Data_Benthos.xlsx
    + Update with TolVal2 for example in Vignette to avoid warning.
* Vignette
    + SITE_TYPE to INDEX_REGION
    + qc.checks
        - Update example chunk to match example in function; package BCGcalc to BioMonTools.

## Changes in Version 0.3.0.9002 (2019-01-14)

* Vignette
    + kable used without library(knitr) in rarify chunk.

## Changes in Version 0.3.0.9001 (2018-12-21)

* Update QC flags input.
* Add metric names to Excel file.
* MetricName to METRIC_NAME in input files.

## Changes in Version 0.3 (2018-12-20)

* New release version.

## Changes in Version 0.2.0.9007 (2018-12-20)

* Metric scoring function, Issue #12.

## Changes in Version 0.2.0.9006 (2018-12-03)

* README.rmd, Issue #11.
    + install_github command comes out as smart quotes if copy and paste from GitHub website.
    + Looks ok in RStudio but fixed to test if that is the issue.

## Changes in Version 0.2.0.9005 (2018-11-26)

* Added MapTaxaObs.

## Changes in Version 0.2.0.9004 (2018-11-26)

* Percent metrics as 0-1.  Issue #9.

## Changes in Version 0.2.0.9003 (2018-11-26)

* Add qc.checks function from BCGcalc package.  Issue #9
    + qc.checks.R
    + Vignette
    + extdata\MetricFlags.xlsx
* Update read_excel with guess_max=10^6 for bio data file.

## Changes in Version 0.2.0.9002 (2018-11-20)

* Update SuperFamily Tipuloidea.
    + Update data and examples in function and vignette

## Changes in Version 0.2.0.9001 (2018-11-20)

* Add Travis CI.

## Changes in Version 0.2.0 (2018-11-20)

* Initial release version.

## Changes in Version 0.1.0.9004 (2018-11-20)

* Add SuperFamily
    + function
    + vignette
    + example data file
* Add metric names files to extdata and doc

## Changes in Version 0.1.0.9003 (2018-11-20)

* markExcluded
    + Add function with examples
* extdata\Data_Benthos.xlsx
    + Updated bad Excluded entries.
        - FALSE to TRUE, n=6
        - TRUE to FALSE, n=1
* DESCRIPTION
    + Update Imports and Suggests
* NEWS
    + Update Planned and Future Updates
* README
    + Tweak language in Usage section.
    + Update icons.
    + Add Travis CI.
* Vignette

## Changes in Version 0.1.0.9002 (2018-11-16)

* extdata\Data_Benthos.xlsx
    + OneDrive messed up version included in package.

## Changes in Version 0.1.0.9001 (2018-11-16)

* Added basic package structure.
    + Folders (data, data-raw, inst/doc, inst/extdata, man, R, vignettes)
    + NEWS, README, LICENSE
* Update DESCRIPTION, NEWS, and README
* Add functions and data
    + rarify
    + metric.values

## Changes in Version 0.1.0 (2018-11-16)

* Initial commit on GitHub.

