---
title: "Library Creation"
subtitle: "BioMonTools"
author: "Erik.Leppo@tetratech.com"
date: "`r Sys.time()`"
output: 
  html_notebook:
    toc: yes
    depth: 3
    toc_float: no
---

Helper code for creating library.

Erik.Leppo@tetratech.com
2017-09-26

# Package

Package related code.

## Vignette

1. Need to run code below to add data to "vignette" directory so it will work.

```{r Vignette_Data, eval=FALSE}

```

2. Knit the vignette in the vignette folder to ensure all of the code works and 
is properly displayed.  This can take 2 to 5 min.

3. Use the code below to "build" the vignette and then use it in the pkg.

```{r Vignette_Build, eval=FALSE}
# generate Vignette
library(BioMonTools)
library(devtools)
devtools::build_vignettes()

# create vignette folder and default file
#devtools::use_vignette("BCGcalc_Vignette")
```
devtools::build() more useful and build_vignettes()

## Create Package
Use the code below to create the package.  Assumes you are in the development package directory

```{r, eval=FALSE}
# Library Name
# Remove all files in "Results" folder
# Triggered here so can run different files
path_results <- file.path("inst", "shiny-examples", "BioMonTools", "results")
fn_results <- list.files(path_results
                         , full.names = TRUE
                         , include.dirs = TRUE
                         , recursive = TRUE)
unlink(fn_results, recursive = TRUE) # to include dir use unlink instead of file.remove
# Copy file to ensure directory not empty
path_shiny <- file.path("inst", "shiny-examples", "BioMonTools")
fn_copy    <- "remove.txt"
path_from  <- file.path(path_shiny, "data", fn_copy)
path_to    <- file.path(path_shiny, "results", fn_copy)
file.copy(path_from, path_to)
# NEWS
# Render then Copy NEWS so picked up in help
rmarkdown::render("NEWS.rmd", "all")
file.copy("NEWS.md", "NEWS", overwrite = TRUE)
file.remove("NEWS.html")
#file.remove("NEWS.md")
#
myPkg <- "BioMonTools"
# Load Library
library(devtools)
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Create Package
# create(myPkg)
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Document, Install, and Reload Library
## Generate Documentation
#setwd(paste0("./",myPkg))  # will fail as already in this directory
devtools::document()
## Install New Package (locally)
#setwd("..") # return to root directory first
# change to build_vignettes = TRUE, use FALSE for speed
devtools::install(".", quick = FALSE, reload = TRUE, build_vignettes = FALSE) 

# blank line above allows code to continue when install asks about package updates 
## Reload library
library(myPkg, character.only = TRUE)
# change wd back to package
#setwd(paste0("./",myPkg))
# Help page
help(package=(myPkg))
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

After creating the package reload it after restarting R within RStudio (Ctrl + Shift + F10).

```{r, eval=FALSE}
# Restart R within RStudio:  Ctrl + Shift + F10
myPkg <- "BioMonTools"
library(myPkg, character.only = TRUE)
help(package=(myPkg))
library(testthat)
#
citation(myPkg)
```

# Package Down
Create pkgdown website.
```{r PkgDown}
pkgdown::build_site()
```


**code coverage**
```{r CodeCoverage}
#library(covr)
covr::report()
```

# other good stuff
```{r}
g <- goodpractice::gp()

s_check <- devtools::spell_check()

```


### Run All Examples
Run all examples at once rather than going to each function individually.

Add "/_test" to gitignore

(Needs tweaking to get it working properly)

```{r AllExamples, eval=FALSE}
#pkg <- "BioMonTools" # from root directory of pkg
#setwd("./_test")
devtools::run_examples()
```

### Data Documentation
Use promptData to get data.r listing for columns.  Short cuts creating documentation.

```{r DocData}
promptData(data_diatom_mmi_qc)
```

## Check

```{r, eval=FALSE}
# Check for errors (or press Cmd + Shift + E in RStudio)
#http://r-pkgs.had.co.nz/check.html
devtools::check()
```

If use RStudio, Build - Check Package, the output is easier to read. 


## Test
Need to use testthat but not to that stage yet.

RStudio, Build - Test Package.


# CI, Badges, Code Coverage, etc
* Continuous Integration with GitHub Actions (TravisCI shut down Dec 2020)
* Code Quality (CodeFactor)
    + CodeFactor.io
        - GitHub login
* Code Coverage (testing)
    + Codecov.io
        - GitHub login
        - After set up CI with GitHub Actions
* Package website (pkgdown and GitHub Action)
    + branch gh-pages, auto updates with each commit
    + Update in settings the location
* Badges; Maintained, Life Cycle, License, Issues, etc
    + https://github.com/GuangchuangYu/badger
        - Commands to create badges

badger package creates the text for pkg readme file.
        
Also
https://shields.io/        
https://github.com/badges/shields
Can make some badges dynamic, static ones below

# issues
[![GitHub issues](https://img.shields.io/github/issues/leppott/MBSStools.svg)](https://GitHub.com/leppott/MBSStools/issues/)
# release
[![GitHub release](https://img.shields.io/github/release/leppott/MBSStools.svg)](https://GitHub.com/leppott/MBSStools/releases/)
# downloads
[![Github all releases](https://img.shields.io/github/downloads/leppott/MBSStools/total.svg)](https://GitHub.com/leppott/MBSStools/releases/)
        
```{r}
pkg_GH <- "leppott/BioMonTools"
pkg_lic <- "MIT"

# GitHub Action, CI
usethis::use_github_actions()
usethis::use_github_actions_badge()
# badger::badge_github_actions(pkg_GH)

# code coverage
usethis::use_github_action("test-coverage")

# GitHub Action, pkgdown
usethis::use_github_action("pkgdown")

# Badges
# lifecycle
badger::badge_lifecycle("stable", "green") # stable/green, maturing/dormant/blue
# maintained
# license
badger::badge_license(pkg_lic)
# code quality
badger::badge_codefactor(pkg_GH)
# code testing coverage
badger::badge_codecov(pkg_GH)
# cmd check
badger::badge_github_actions(pkg_GH)


```




# Other Stuff

```{r, eval=FALSE}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Upload to Github via GitHub Desktop utility
# 0. download from web via "clone or download" via "Open in Desktop" (GitHub Desktop) if not already in GitHub Desktop
# 1. Make changes in download/clone folder. (done above)
# 3. Open GH Desktop commit changes then sync.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# install from GitHub (via devtools)
devtools::install_github(paste0("leppott/",myLibrary))
#



#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# remove installed packages (if needed for troubleshooting)
search() # find
#detach(3) # remove by number
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~


# to build package
#https://thepoliticalmethodologist.com/2014/08/14/building-and-maintaining-r-packages-with-devtools-and-roxygen2/
# To build the package as a compressed file in your working directory, run build(current.code, path=getwd()).

# to save internal data for examples
# example
#http://r-pkgs.had.co.nz/data.html#data-sysdata
# have to be at root directory (above package)
#devtools::use_data(NV.predictors,NV.bugs,pkg="MMIcalcNV",internal=TRUE,overwrite=TRUE)
## verify with data()

# To save RMD files
# http://stackoverflow.com/questions/30377213/how-to-include-rmarkdown-file-in-r-package
# /pkg/inst/rmd/
# system.file("rmd/file.Rmd", package="packagename")
#
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#https://hilaryparker.com/2014/04/29/writing-an-r-package-from-scratch/
# Create Package
# create(myLibrary)

```


