---
title: "Library Creation"
subtitle: "BioMonTools"
author: "Erik.Leppo@tetratech.com"
date: "`r Sys.time()`"
output: 
  html_notebook:
    toc: yes
    depth: 3
    toc_float: no
---

Helper code for creating library.

Erik.Leppo@tetratech.com
2017-09-26

# Package

Package related code.

## Vignette

1. Need to run code below to add data to "vignette" directory so it will work.

```{r Vignette_Data, eval=FALSE}

```

2. Knit the vignette in the vignette folder to ensure all of the code works and 
is properly displayed.  This can take 2 to 5 min.

3. Use the code below to "build" the vignette and then use it in the pkg.

```{r Vignette_Build, eval=FALSE}
# generate Vignette
library(BioMonTools)
library(devtools)
devtools::build_vignettes()

# create vignette folder and default file
#devtools::use_vignette("BCGcalc_Vignette")
```
devtools::build() more useful and build_vignettes()

## Create Package
Use the code below to create the package.  Assumes you are in the development package directory

```{r, eval=FALSE}
# Library Name
# NEWS
# Render then Copy NEWS so picked up in help
rmarkdown::render("NEWS.rmd", "all")
file.copy("NEWS.md", "NEWS", overwrite = TRUE)
file.remove("NEWS.html")
file.remove("NEWS.md")
#
myPkg <- "BioMonTools"
# Load Library
library(devtools)
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Create Package
# create(myPkg)
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Document, Install, and Reload Library
## Generate Documentation
setwd(paste0("./",myPkg))  # will fail as already in this directory
devtools::document()
## Install New Package (locally)
setwd("..") # return to root directory first
devtools::install(myPkg)
## Reload library
library(myPkg, character.only = TRUE)
# change wd back to package
setwd(paste0("./",myPkg))
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

After creating the package reload it after restarting R within RStudio (Ctrl + Shift + F10).

```{r, eval=FALSE}
# Restart R within RStudio:  Ctrl + Shift + F10
myPkg <- "BioMonTools"
library(BCGcalc)
help(package=(myPkg))
#
citation(myPkg)
```

### Run All Examples
Run all examples at once rather than going to each function individually.

Add "/_test" to gitignore

(Needs tweaking to get it working properly)

```{r AllExamples, eval=FALSE}
pkg <- "BCGcalc"
setwd("./_test")
devtools::run_examples(pkg=pkg)
```


## Including Data
All data should have already been created.

## Check

```{r, eval=FALSE}
# Check for errors (or press Cmd + Shift + E in RStudio)
#http://r-pkgs.had.co.nz/check.html
devtools::check()
```

If use RStudio, Build - Check Package, the output is easier to read. 


## Test
Need to use testthat but not to that stage yet.

RStudio, Build - Test Package.


# Other Stuff

```{r, eval=FALSE}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Upload to Github via GitHub Desktop utility
# 0. download from web via "clone or download" via "Open in Desktop" (GitHub Desktop) if not already in GitHub Desktop
# 1. Make changes in download/clone folder. (done above)
# 3. Open GH Desktop commit changes then sync.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# install from GitHub (via devtools)
devtools::install_github(paste0("leppott/",myLibrary))
#



#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# remove installed packages (if needed for troubleshooting)
search() # find
#detach(3) # remove by number
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~


# to build package
#https://thepoliticalmethodologist.com/2014/08/14/building-and-maintaining-r-packages-with-devtools-and-roxygen2/
# To build the package as a compressed file in your working directory, run build(current.code, path=getwd()).

# to save internal data for examples
# example
#http://r-pkgs.had.co.nz/data.html#data-sysdata
# have to be at root directory (above package)
#devtools::use_data(NV.predictors,NV.bugs,pkg="MMIcalcNV",internal=TRUE,overwrite=TRUE)
## verify with data()

# To save RMD files
# http://stackoverflow.com/questions/30377213/how-to-include-rmarkdown-file-in-r-package
# /pkg/inst/rmd/
# system.file("rmd/file.Rmd", package="packagename")
#
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#https://hilaryparker.com/2014/04/29/writing-an-r-package-from-scratch/
# Create Package
# create(myLibrary)

```


